Exploiting First-Class Arrays in Fortran for Accelerator Programming

### 2010-12-03 opencl/papers/foropencl.tex
edits

This paper presents a compiler-level approach for targeting a single
program to multiple fundamentally different low level execution and
programming models while allowing the application programmer to adopt
a single high level programming model.

@@

A data parallel programming model is suited for accelerators....

One possibility is to adopt a new language like CUDA or Chapel...
Another possibility is to use a mature language with OpenMP directives
or write OpenCL kernels.

The problem is that many of the language choices expose too much detail
about the machine architecture.  This is particularly true of CUDA and OpenCL.

At this point in time it is not really possible to write once and
run efficiently on the wide variety of computer platforms we have
available.

@@

>>>>

This paper presents a compiler-level approach for targeting a single
program to multiple fundamentally different low level execution and
programming models while allowing the application programmer to adopt
a single high level programming model.  We suggest that features of
Fortran 90 for data parallel programming are well suited to automatic
transformation to generate code specifically tuned for different
low-level programming models such as OpenCL and CUDA.

@@

A data parallel programming model is well suited to emerging architectures
that are based on either vector processing or massively parallel collections
of simple cores.  A popular trend for programming systems such as GPU
accelerators is the CUDA language from NVidia, or new languages like Chapel
from Cray.  A less intrusive approach that is well suited to legacy 
applications and languages are language extensions or libraries that allow
programmers to avoid adopting entirely new languages.  The recent OpenCL
specification is intended to support this programming model, as are
directive-based methods such as OpenMP or the ``Accelerate'' product from
the Portland Group.

The problem with many of these choices is that they expose too much detail
about the machine architecture to the programmer.  
This is particularly true of CUDA and OpenCL.  In CUDA, programmers must
adapt their codes to fit the threading model used by NVidia GPUs, while
OpenCL requires programmers to provide specially tuned versions of
their code for different classes of machine.  In both cases, the programmer
is also responsible for explicitly managing memory, including staging of
data back and forth from the host CPU and the accelerator device memory.
While these models have been attractive as a method for early adopters
to utilize these new architectures, they are less attractive to programmers
who do not have the time or resources to manually port their code to every
new architecture and programming model that emerges.

At this point in time it is not really possible to write once and
run efficiently on the wide variety of computer platforms we have
available.  For some classes of applications, we believe that this goal
is possible using language constructs already present in a popular
mainstream scientific programming language -- Fortran.

\subsection{Approach}

@@

\section{Junk}

\cite{chamberlain04zpl}

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,foropencl}
